---
title: "Big Mart Sales Prediction"
output: html_notebook
---

## Sales Prediction for Big Mart Outlets 

BigMart sales data for 1559 products across 10 stores in different cities. certain variables of each product and store have been defined. The aim is to build a predictive model and predict the sales of each product at a particular outlet.


| Variable                  	| Description                                                                                 	|
|---------------------------	|---------------------------------------------------------------------------------------------	|
| Item_Identifier           	| Unique product ID                                                                           	|
| Item_Weight               	| Weight of product                                                                           	|
| Item_Fat_Content          	| Whether the product is low fat or not                                                       	|
| Item_Visibility           	| The % of total display area of all products in a store allocated to the particular product  	|
| Item_Type                 	| The category to which the product belongs                                                   	|
| Item_MRP                  	| Maximum Retail Price (list price) of the product                                            	|
| Outlet_Identifier         	| Unique store ID                                                                             	|
| Outlet_Establishment_Year 	| The year in which store was established                                                     	|
| Outlet_Size               	| The size of the store in terms of ground area covered                                       	|
| Outlet_Location_Type      	| The type of city in which the store is located                                              	|
| Outlet_Type               	| Whether the outlet is just a grocery store or some sort of supermarket                      	|
| Item_Outlet_Sales         	| Sales of the product in the particular store. This is the outcome variable to be predicted. 	|



### Loading Libraries
```{r}
library(readr)
```


### Load Datasets

```{r}
train = read_csv("/cloud/project/Data/Big Mart Sales.csv")
head(train)
```


```{r}
# check dimensions ( number of row & columns) in data set

dim(train)
```

```{r}
#check the variables and their types in train
str(train)
```


```{r}
# Check if data has missing values
table(is.na(train))

summary(is.na(train))

colSums(is.na(train))
```

### Summary of Train data

```{r}
summary(train)
```

### Graphical Representation of Variables

```{r}
#Item_Visibility ~ Item_Outlet_Sales
ggplot(train, aes(x= Item_Visibility, y = Item_Outlet_Sales)) + 
  geom_point(size = 2, color="navy") + 
  xlab("Item Visibility") + ylab("Item Outlet Sales") + 
  ggtitle("Item Visibility vs Item Outlet Sales") + 
  theme_linedraw()
```

```{r}
#Outlets vs Total Sales
ggplot(train, aes(Outlet_Identifier, Item_Outlet_Sales)) + 
  geom_bar(stat = "identity", color = "purple") +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black"))  +
  ggtitle("Outlets vs Total Sales") + theme_bw()
```


```{r}
# Item Type vs Sales
ggplot(train, aes(Item_Type, Item_Outlet_Sales)) + 
  geom_bar( stat = "identity",color = "blue") +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "navy")) + 
  xlab("Item Type") + ylab("Item Outlet Sales")+
  ggtitle("Item Type vs Sales") +
  theme_void()


```


```{r}
#Item Type vs Item MRP
ggplot(train, aes(Item_Type, Item_MRP)) +
  geom_boxplot() +
  ggtitle("Box Plot") + theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "red")) + 
  xlab("Item Type") + ylab("Item MRP") + 
  ggtitle("Item Type vs Item MRP")


```


### Importing Test data
```{r}
test = read_csv("/cloud/project/Data/test.csv")
head(test)
```


```{r}
#Checking Dim of test and train data
dim(train)
dim(test)
```

```{r}
colnames(train)
print(".........................................................")
colnames(test)
```

```{r}
# combine the data sets.
test$Item_Outlet_Sales <-  1

combi <- rbind(train, test)
dim(combi)
```

### Impute missing value 

```{r}
#Impute missing value by median
combi$Item_Weight[is.na(combi$Item_Weight)] <- median(combi$Item_Weight, na.rm = TRUE)
combi$Item_Weight[is.na(combi$Item_Weight)]
table(is.na(combi$Item_Weight))

```

In the graph above, the item visibility has zero value also, which is practically not feasible. Hence, we’ll consider it as a missing value so we made the imputation using median.
```{r}
combi$Item_Visibility <- ifelse(combi$Item_Visibility == 0,
                           median(combi$Item_Visibility), combi$Item_Visibility)
```



```{r}
#Item_Visibility ~ Item_Outlet_Sales
ggplot(combi, aes(x= Item_Visibility, y = Item_Outlet_Sales)) + 
  geom_point(size = 2, color="navy") + 
  xlab("Item Visibility") + ylab("Item Outlet Sales") + 
  ggtitle("Item Visibility vs Item Outlet Sales") + 
  theme_linedraw()
```


## Data Manipulation in R












































