---
title: "Big Mart Sales Prediction"
output: html_notebook
---

## Sales Prediction for Big Mart Outlets 

BigMart sales data for 1559 products across 10 stores in different cities. certain variables of each product and store have been defined. The aim is to build a predictive model and predict the sales of each product at a particular outlet.


| Variable                  	| Description                                                                                 	|
|---------------------------	|---------------------------------------------------------------------------------------------	|
| Item_Identifier           	| Unique product ID                                                                           	|
| Item_Weight               	| Weight of product                                                                           	|
| Item_Fat_Content          	| Whether the product is low fat or not                                                       	|
| Item_Visibility           	| The % of total display area of all products in a store allocated to the particular product  	|
| Item_Type                 	| The category to which the product belongs                                                   	|
| Item_MRP                  	| Maximum Retail Price (list price) of the product                                            	|
| Outlet_Identifier         	| Unique store ID                                                                             	|
| Outlet_Establishment_Year 	| The year in which store was established                                                     	|
| Outlet_Size               	| The size of the store in terms of ground area covered                                       	|
| Outlet_Location_Type      	| The type of city in which the store is located                                              	|
| Outlet_Type               	| Whether the outlet is just a grocery store or some sort of supermarket                      	|
| Item_Outlet_Sales         	| Sales of the product in the particular store. This is the outcome variable to be predicted. 	|



### Loading Libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(caret)
library(caretEnsemble)
library(VIM)
library(gridExtra)
library(cowplot)
library(data.table)
library(corrplot)
library(e1071)
library(xgboost)
```


### Load Datasets

```{r}
train = read_csv("/cloud/project/Data/BigMartSales.csv")
head(train)
```


```{r}
# check dimensions ( number of row & columns) in data set

dim(train)
```

```{r}
#check the variables and their types in train
str(train)
```


```{r}
# Check if data has missing values
table(is.na(train))

summary(is.na(train))

colSums(is.na(train))
```

### Summary of Train data

```{r}
summary(train)
```

### Graphical Representation of Variables

```{r}
#Item_Visibility ~ Item_Outlet_Sales
ggplot(train, aes(x= Item_Visibility, y = Item_Outlet_Sales)) + 
  geom_point(size = 2, color="navy") + 
  xlab("Item Visibility") + ylab("Item Outlet Sales") + 
  ggtitle("Item Visibility vs Item Outlet Sales") + 
  theme_linedraw()
```

```{r}
#Outlets vs Total Sales
ggplot(train, aes(Outlet_Identifier, Item_Outlet_Sales)) + 
  geom_bar(stat = "identity", color = "purple") +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black"))  +
  ggtitle("Outlets vs Total Sales") + theme_bw()
```


```{r}
# Item Type vs Sales
ggplot(train, aes(Item_Type, Item_Outlet_Sales)) + 
  geom_bar( stat = "identity",color = "blue") +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "navy")) + 
  xlab("Item Type") + ylab("Item Outlet Sales")+
  ggtitle("Item Type vs Sales") +
  theme_void()


```


```{r}
#Item Type vs Item MRP
ggplot(train, aes(Item_Type, Item_MRP)) +
  geom_boxplot() +
  ggtitle("Box Plot") + theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "red")) + 
  xlab("Item Type") + ylab("Item MRP") + 
  ggtitle("Item Type vs Item MRP")


```


### Importing Test data
```{r}
test = read_csv("/cloud/project/Data/test.csv")
head(test)
```


```{r}
#Checking Dim of test and train data
dim(train)
dim(test)
```

```{r}
colnames(train)
print(".........................................................")
colnames(test)
```

```{r}
# combine the data sets.
test$Item_Outlet_Sales <-  1

combi <- rbind(train, test)
dim(combi)
```

### Impute missing value 

```{r}
#Impute missing value by median
combi$Item_Weight[is.na(combi$Item_Weight)] <- median(combi$Item_Weight, na.rm = TRUE)
combi$Item_Weight[is.na(combi$Item_Weight)]
table(is.na(combi$Item_Weight))

```

In the graph above, the item visibility has zero value also, which is practically not feasible. Hence, we’ll consider it as a missing value so we made the imputation using median.
```{r}
combi$Item_Visibility <- ifelse(combi$Item_Visibility == 0,
                           median(combi$Item_Visibility), combi$Item_Visibility)
```



```{r}
#Item_Visibility ~ Item_Outlet_Sales
ggplot(combi, aes(x= Item_Visibility, y = Item_Outlet_Sales)) + 
  geom_point(size = 2, color="navy") + 
  xlab("Item Visibility") + ylab("Item Outlet Sales") + 
  ggtitle("Item Visibility vs Item Outlet Sales") + 
  theme_linedraw()
```


## Data Manipulation in R



### Load Datasets

```{r}
big_mart = read_csv("/cloud/project/Data/BigMartSales.csv")
head(train)

```

```{r}
glimpse(big_mart)
```

```{r}
summary(big_mart)
```

## Manipulating Dataset



```{r}
unique(big_mart$Item_Fat_Content)
```
**Cleaning Item_Fat_Content Variable**
```{r}
#Transforming "low fat" and "LF" to "Low Fat"
index <- which(big_mart$Item_Fat_Content == "LF" | 
                 big_mart$Item_Fat_Content == "low fat")

big_mart[index, "Item_Fat_Content"] <- "Low Fat"

#Transforming "reg" to "Regular
index2 <- which(big_mart$Item_Fat_Content == "reg")

big_mart[index2, "Item_Fat_Content"] <- "Regular"

#Dropping Unused Levels
big_mart$Item_Fat_Content <- factor(big_mart$Item_Fat_Content)

unique(big_mart$Item_Fat_Content)
```
Imputing Missing Values
```{r}
#Using kNN imputation for missing values
big_mart_imputed <- kNN(big_mart)
big_mart_imputed <- big_mart_imputed %>% 
    select(Item_Identifier:Item_Outlet_Sales)

summary(big_mart_imputed)

summary(is.na(big_mart_imputed))
```

### Impute Values for Outlet_Size
```{r}
#Outlet Identifier by Outlet Size Table
table(big_mart_imputed$Outlet_Identifier, big_mart_imputed$Outlet_Size)

```
```{r}
# Outlet Identifier by Outlet_Type Table
table(big_mart_imputed$Outlet_Identifier, big_mart_imputed$Outlet_Type)

```

```{r}
# Outlet Type by Outlet Size Table
table(big_mart$Outlet_Type, big_mart_imputed$Outlet_Size)

```

```{r}
# Imputing Small for OUT010 Location
index3 <- which(big_mart_imputed$Outlet_Identifier == "OUT010")
big_mart_imputed[index3, "Outlet_Size"] <- "Small"

#Imputing Small for OUT017 Location
index4 <- which(big_mart_imputed$Outlet_Identifier == "OUT017")
big_mart_imputed[index4, "Outlet_Size"] <- "Medium"

# Imputing Medium for OUT045 Location
index5 <- which(big_mart_imputed$Outlet_Identifier == "OUT045")
big_mart_imputed[index5, "Outlet_Size"] <- "Medium"

# Dropping Unused Levels for Outlet Size Variable
big_mart_imputed$Outlet_Size <- factor(big_mart_imputed$Outlet_Size)

#Summary Cleaned Dataset
summary(big_mart_imputed)
```

# Visualizing Data
```{r}
# Item Outlet Sales Histogram

ggplot(big_mart_imputed, aes(x=Item_Outlet_Sales)) +
  geom_histogram(binwidth = 200,colour = c("blue")) +
  labs(title = "Item Outlet Sales Histogram", x = "Item Outlet Sales")+
  theme_light()

```

```{r}
# Item Outlet Sales Histogram by Outlet Identifier

ggplot(big_mart_imputed, aes(x=Item_Outlet_Sales, fill = Outlet_Identifier)) +
  geom_histogram(binwidth = 200) +
  facet_wrap(~ Outlet_Identifier) +
  labs(title = "Item Outlet Sales Histogram", x = "Item Outlet Sales") +
  theme_minimal()
```

```{r}
# Sales by Outlet Identifier

ggplot(big_mart_imputed, aes(x = Outlet_Identifier,y = Item_Outlet_Sales)) +
  geom_boxplot() +
  labs(title = "Sales by Outlet Identifier", x = "Outlet Identifier", y = "Item Outlet Sales") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# Item Outlet Sales by Item MRP and Outlet Identifier
ggplot(big_mart_imputed, aes(x = Item_MRP,y = Item_Outlet_Sales)) +
  geom_bin2d() +
  facet_wrap(~ Outlet_Identifier) +
  labs(title = "Item Outlet Sales by Item MRP and Outlet Identifier",
       x = "Item MRP",
       y = "Item Outlet Sales")
```

### Median Sales by Location
```{r}

sales = big_mart_imputed %>%
  group_by(Outlet_Identifier) %>%
  summarize(median_sales = median(Item_Outlet_Sales))%>%
  arrange(desc(median_sales))

sales
```


```{r}
# Correlation of Item Outlet Sales and Item MRP
cor(big_mart_imputed$Item_MRP, big_mart_imputed$Item_Outlet_Sales)
```

## Modeling
*Linear Regression
*Lasso Regression
*Ridge Regression
*RandomForest
*XGBoost

```{r}
head(train)
head(test)
dim(train)
dim(test)
dim(combi)
```



```{r}
#plotting sales histogram.
ggplot(train) + geom_histogram(aes(train$Item_Outlet_Sales), binwidth = 100, fill = "darkgreen") +
  xlab("Item_Outlet_Sales")
```

```{r}
#Independent Variables (numeric variables)
p1 = ggplot(combi) + geom_histogram(aes(Item_Weight), binwidth = 0.5, fill = "blue")
p2 = ggplot(combi) + geom_histogram(aes(Item_Visibility), binwidth = 0.005, fill = "blue")
p3 = ggplot(combi) + geom_histogram(aes(Item_MRP), binwidth = 1, fill = "blue")
plot_grid(p1, p2, p3, nrow = 3) # plot_grid() from cowplot package
```

```{r}
# Independent Variables (categorical variables)
p4=ggplot(combi %>% group_by(Item_Fat_Content) %>% summarise(Count = n())) + 
  geom_bar(aes(Item_Fat_Content, Count), stat = "identity", fill = "coral1")
combi$Item_Fat_Content[combi$Item_Fat_Content == "LF"] = "Low Fat"
combi$Item_Fat_Content[combi$Item_Fat_Content == "low fat"] = "Low Fat"
combi$Item_Fat_Content[combi$Item_Fat_Content == "reg"] = "Regular"
p5=ggplot(combi %>% group_by(Item_Fat_Content) %>% summarise(Count = n())) + 
  geom_bar(aes(Item_Fat_Content, Count), stat = "identity", fill = "coral1")

plot_grid(p4,p5,nrow = 1)
```

```{r}
p6 = ggplot(combi %>% group_by(Item_Type) %>% summarise(Count = n())) + 
  geom_bar(aes(Item_Type, Count), stat = "identity", fill = "coral1") +
  xlab("") +
  geom_label(aes(Item_Type, Count, label = Count), vjust = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Item_Type")
p7 = ggplot(combi %>% group_by(Outlet_Identifier) %>% summarise(Count = n())) + 
  geom_bar(aes(Outlet_Identifier, Count), stat = "identity", fill = "coral1") +
  geom_label(aes(Outlet_Identifier, Count, label = Count), vjust = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p8 = ggplot(combi %>% group_by(Outlet_Size) %>% summarise(Count = n())) + 
  geom_bar(aes(Outlet_Size, Count), stat = "identity", fill = "coral1") +
  geom_label(aes(Outlet_Size, Count, label = Count), vjust = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
second_row = plot_grid(p7, p8, nrow = 1)
plot_grid(p6, second_row, ncol = 1)
```


```{r}
p9 = ggplot(combi %>% group_by(Outlet_Establishment_Year) %>% summarise(Count = n())) + 
  geom_bar(aes(factor(Outlet_Establishment_Year), Count), stat = "identity", fill = "coral1") +
  geom_label(aes(factor(Outlet_Establishment_Year), Count, label = Count), vjust = 0.5) +
  xlab("Outlet_Establishment_Year") +
  theme(axis.text.x = element_text(size = 8.5))
p10 = ggplot(combi %>% group_by(Outlet_Type) %>% summarise(Count = n())) + 
  geom_bar(aes(Outlet_Type, Count), stat = "identity", fill = "coral1") +
  geom_label(aes(factor(Outlet_Type), Count, label = Count), vjust = 0.5) +
  theme(axis.text.x = element_text(size = 8.5))
plot_grid(p9, p10, ncol = 2)
```


```{r}
#Split train data
train = combi[1:nrow(train),]
head(train)
dim(train)
```
### Bivariate Analysis
```{r}
p9 = ggplot(train) + geom_point(aes(Item_Weight, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +
     theme(axis.title = element_text(size = 8.5))
p10 = ggplot(train) + geom_point(aes(Item_Visibility, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +
      theme(axis.title = element_text(size = 8.5))
p11 = ggplot(train) + geom_point(aes(Item_MRP, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +
      theme(axis.title = element_text(size = 8.5))
second_row_2 = plot_grid(p10, p11, ncol = 2)
plot_grid(p9, second_row_2, nrow = 2)
```


```{r}
p12 = ggplot(train) + geom_violin(aes(Item_Type, Item_Outlet_Sales), fill = "magenta") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 6),
        axis.title = element_text(size = 8.5))
p13 = ggplot(train) + geom_violin(aes(Item_Fat_Content, Item_Outlet_Sales), fill = "magenta") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8.5))
p14 = ggplot(train) + geom_violin(aes(Outlet_Identifier, Item_Outlet_Sales), fill = "magenta") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8.5))
second_row_3 = plot_grid(p13, p14, ncol = 2)
plot_grid(p12, second_row_3, ncol = 1)
```


```{r}
ggplot(train) + geom_violin(aes(Outlet_Size, Item_Outlet_Sales), fill = "magenta")
```


```{r}
p15 = ggplot(train) + geom_violin(aes(Outlet_Location_Type, Item_Outlet_Sales), fill = "magenta")
p16 = ggplot(train) + geom_violin(aes(Outlet_Type, Item_Outlet_Sales), fill = "magenta")
plot_grid(p15, p16, ncol = 1)
```


### Missing Value Treatment

```{r}
colSums(is.na(combi))
```

```{r}
#Replacing 0’s in Item_Visibility variable
zero_index = which(combi$Item_Visibility == 0)
for(i in zero_index){
  
  item = combi$Item_Identifier[i]
  combi$Item_Visibility[i] = mean(combi$Item_Visibility[combi$Item_Identifier == item], na.rm = T)
  
}
```

```{r}
# create a new feature 'Item_Type_new' 
perishable = c("Breads", "Breakfast", "Dairy", "Fruits and Vegetables", "Meat", "Seafood")
non_perishable = c("Baking Goods", "Canned", "Frozen Foods", "Hard Drinks", "Health and Hygiene","Household", "Soft Drinks")
combi$Item_Type_new0 = ifelse(combi$Item_Type %in% perishable, "perishable",
                               ifelse(combi$Item_Type %in% non_perishable, "non_perishable", "not_sure"))
```


```{r}
table(combi$Item_Type, substr(combi$Item_Identifier, 1, 2))
```

```{r}
combi$Item_category = substr(combi$Item_Identifier, 1, 2)
combi$Item_Fat_Content[combi$Item_category == "NC"] = "Non-Edible"
# years of operation of outlets
combi$Outlet_Years = 2013 - as.numeric(combi$Outlet_Establishment_Year)
combi$Outlet_Establishment_Year = as.factor(combi$Outlet_Establishment_Year)
# Price per unit weight
combi$price_per_unit_wt = combi$Item_MRP/combi$Item_Weight
```


```{r}
Item_MRP_clusters = kmeans(combi$Item_MRP, centers = 4)
table(Item_MRP_clusters$cluster) # display no. of observations in each cluster
```

```{r}
combi$Item_MRP_clusters = as.factor(Item_MRP_clusters$cluster)
```


## Encoding Categorical Variables

### Label encoding for the categorical variables

We will label encode *Outlet_Size* and *Outlet_Location_Type* as these are ordinal variables.

```{r}
combi$Outlet_Size_num = ifelse(combi$Outlet_Size == "Small", 0,
                                 ifelse(combi$Outlet_Size == "Medium", 1, 2))
combi$Outlet_Location_Type_num = ifelse(combi$Outlet_Location_Type == "Tier 3", 0,
                                          ifelse(combi$Outlet_Location_Type == "Tier 2", 1, 2))
# removing categorical variables after label encoding
#combi[, c("Outlet_Size", "Outlet_Location_Type") := NULL]
```


### One hot encoding for the categorical variable
```{r}
ohe = dummyVars("~.", data = combi[,-c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")], fullRank = T)
ohe_df = data.table(predict(ohe, combi[,-c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")]))
combi = cbind(combi[,"Item_Identifier"], ohe_df)
```



















































