---
title: "Novel Corona Virus 2019 Analysis"
output: html_notebook
---

# Novel Corona Virus 2019 Dataset
### Day level information on covid-19 affected cases



This dataset has daily level information on the number of affected cases, deaths and recovery from 2019 novel coronavirus. Please note that this is a time series data and so the number of cases on any given day is the cumulative number.

The data is available from 22 Jan, 2020.

## Variable Description
* Sno - Serial number
* ObservationDate - Date of the observation in MM/DD/YYYY
* Province/State - Province or state of the observation (Could be empty when missing)
* Country/Region - Country of observation
* Last Update - Time in UTC at which the row is updated for the given province or country. (Not standardised and so please clean before using it)
* Confirmed - Cumulative number of confirmed cases till that date
* Deaths - Cumulative number of of deaths till that date
* Recovered - Cumulative number of recovered cases till that date


## Libraries
```{r}
#Import Libraries
library(readr)

```

## Reading Dataset
```{r}
#Import dataset
covid_data <- read_csv("Data/covid_19_data.csv")
head(covid_data)
```


```{r}
str(covid_data)
```

```{r}
# Number of columns 
length(covid_data)
# Dimension of data(rows,columns)
dim(covid_data)
```

```{r}
summary(covid_data)
```

```{r}
unique(covid_data$ObservationDate)
unique(covid_data$`Province/State`)
unique(covid_data$`Country/Region`)
```

```{r}
#Plots to explore data
plot(covid_data$Confirmed)
barplot(covid_data$Confirmed)
```
```{r}
#Checking Missing Values
is.na(covid_data)

summary(is.na(covid_data))
```

```{r}
#Identifying less than zero values

any(covid_data$Confirmed<=0)
any(covid_data$Deaths<=0)
any(covid_data$Recovered<=0)
```

















