---
title: "Day 14 - Data Visualization in R"
author: "MAHESH DIVAKARAN"
date: "2/5/2021"
output: html_document
---

# Data Visualization in R

* With the help of visualization techniques, humans can easily gain insights about the hidden patterns in data which might otherwise be neglected. 
* Using data visualization, one can work with large datasets to efficiently obtain key insights about it. 

### R Visualization Packages

* r-base
* ggplot2
* plotly
* shiny
* googleViz
* geofacet
* leaflet
* ggiraph
* tidyquant
* ....

* **Standard Graphics**
* Scatterplots
* Boxplots
* Piecharts
* Barplots etc.

* **Statistical Graphic**
* Data
* Aesthetic Mappings
* Geometric Objects
* Statistical Transformations
* Scales
* Coordinates system
* Faceting
* Wordmaps
* Correlations 
* etc...

### R - Pie Charts

`pie(x, labels, radius, main, col, clockwise)`

* `x` is a vector containing the numeric values used in the pie chart.

* `labels` is used to give description to the slices.

* `radius` indicates the radius of the circle of the pie chart.(value between −1 and +1).

* `main` indicates the title of the chart.

* `col` indicates the color palette.

* `clockwise` is a logical value indicating if the slices are drawn clockwise or anti clockwise.

```{r}
# Create data for the graph.
x <- c(21, 62, 10, 53)
labels <- c("London", "New York", "Singapore", "Mumbai")

# Plot the chart.
pie(x,labels, main = "Pie Chart", col = rainbow(5))

```

```{r}
# Create data for the graph.
x <- c(21, 62, 10, 53)
labels <- c("London", "New York", "Singapore", "Mumbai")

# Give the chart file a name.
png(file = "city.png")

# Plot the chart.
pie(x,labels, main = "Pie Chart", col = rainbow(5))

# Save the file.
dev.off()
```


### Slice Percentages and Chart Legend

```{r}

piepercent<- round(100*x/sum(x), 1)

# Plot the chart.
pie(x, labels = piepercent, main = "City pie chart",col = rainbow(length(x)))
legend("topright", c("London","New York","Singapore","Mumbai"), cex = 0.8,
   fill = rainbow(length(x)))
```


### 3D Pie Chart

```{r}
# Get the library.
library(plotrix)


# Give the chart file a name.
png(file = "3d_pie_chart.jpg")

# Plot the chart.
pie3D(x,labels = labels,explode = 0.1, main = "Pie Chart of Countries ")

# Save the file.
dev.off()
```
## R - Bar Charts

`barplot(H,xlab,ylab,main, names.arg,col)`

* `H` is a vector or matrix containing numeric values used in bar chart.
* `xlab` is the label for x axis.
* `ylab` is the label for y axis.
* `main` is the title of the bar chart.
* `names.arg` is a vector of names appearing under each bar.
* `col` is used to give colors to the bars in the graph.

```{r}
# Create the data for the chart
H <- c(7,12,28,3,41)
label <- c("Mar","Apr","May","Jun","Jul")

# Plot the bar chart 
barplot(H, names.arg=label ,xlab="Month", ylab="Revenue", col="blue", main="Revenue chart",border="red")



# Give the chart file a name
png(file = "barchart_months_revenue.png")

# Plot the bar chart 
barplot(H,names.arg=label,xlab="Month",ylab="Revenue",col="blue",main="Revenue chart",border="red")

# Save the file
dev.off()

```
### Group Bar Chart and Stacked Bar Chart

```{r}
# Create the input vectors.
colors <- c("green","orange","brown")
months <- c("Mar","Apr","May","Jun","Jul")
regions <- c("East","West","North")

# Create the matrix of the values.
Values <- matrix(c(2,9,3,11,9,4,8,7,3,12,5,2,8,10,11), nrow = 3, ncol = 5, byrow = TRUE)
Values

# Give the chart file a name
png(file = "barchart_stacked.png")

# Create the bar chart
barplot(Values, main = "Total revenue", names.arg = months, xlab = "month", ylab = "revenue", col = colors)

# Add the legend to the chart
legend("topleft", regions, cex = 1.3, fill = colors)

# Save the file
dev.off()

```

## R - Histograms

`hist(v,main,xlab,xlim,ylim,breaks,col,border)`

* `v` is a vector containing numeric values used in histogram.

* `main` indicates title of the chart.

* `col` is used to set color of the bars.

* `border` is used to set border color of each bar.

* `xlab` is used to give description of x-axis.

* `xlim` is used to specify the range of values on the x-axis.

* `ylim` is used to specify the range of values on the y-axis.

* `breaks` is used to mention the width of each bar.

```{r}
# Create data for the graph.
v <-  c(9,13,21,8,36,22,12,41,31,33,19)


# Create the histogram.
hist(v,xlab = "Weight",col = "red",border = "blue", breaks = 9)
```



### Range of X and Y values

```{r}
# Create data for the graph.
v <- c(9,13,21,8,36,22,12,41,31,33,19)

# Give the chart file a name.
png(file = "histogram_lim_breaks.png")

# Create the histogram.
hist(v,xlab = "Weight",col = "green",border = "red", xlim = c(0,40), ylim = c(0,5),
   breaks = 5)

# Save the file.
dev.off()
```
## R - Line Graphs

`plot(v,type,col,xlab,ylab)`

* `v` is a vector containing the numeric values.

* `type` takes the value "`p`" to draw only the points, "`l`" to draw only the lines and "`o`" to draw both points and lines.

* `xlab` is the label for x axis.

* `ylab` is the label for y axis.

* `main` is the Title of the chart.

* `col` is used to give colors to both the points and lines.

```{r}
# Create the data for the chart.
v <- c(7,12,28,3,41)


# Plot the bar chart. 
plot(v,type = "o")
plot(v,type = "l")
```

```{r}
# Plot the bar chart.
plot(v,type = "o", col = "red", xlab = "Month", ylab = "Rain fall",
   main = "Rain fall chart")
```


### Multiple Line plots

```{r}
# Create the data for the chart.
v <- c(7,12,28,3,41)
t <- c(14,7,6,19,3)
z <- c(25,1,23,20,31)

# Give the chart file a name.
#png(file = "line_chart_2_lines.jpg")

# Plot the bar chart.
plot(v,type = "o",col = "red", xlab = "Month", ylab = "Rain fall", 
   main = "Rain fall chart")

lines(t, type = "o", col = "blue")
lines(z, type = "o", col = "green")

# Save the file.
#dev.off()
```

## R - Scatterplots

`plot(x, y, main, xlab, ylab, xlim, ylim, axes)`

```{r}
input <- mtcars[,c('wt','mpg')]
print(head(input))
```

```{r}
# Plot the chart
plot(x = input$wt,y = input$mpg)
```

```{r}
# Plot the chart
plot(x = input$wt,y = input$mpg,
   xlab = "Weight",
   ylab = "Milage",
   xlim = c(2.5,5),
   ylim = c(15,30),		 
   main = "Weight vs Milage",
   col = "red"
)
```

## Data Visualization in R using ggplot2


```{r}
library(ggplot2)

head(mtcars)
# Basic barplot
p <- ggplot(mtcars, aes(x=factor(cyl)))+
  geom_bar(stat="count", width=0.5, fill="steelblue") +
  theme_minimal()

p
```

```{r}
# Change barplot line colors by groups
p <- ggplot(mtcars, aes(x=factor(cyl), y=mpg, color=cyl)) +
  geom_bar(stat="identity", fill="white") +
  theme_classic()
p
```

```{r}
# Change bar fill colors to blues
p <- p+scale_fill_brewer(palette="Blues")
p + theme(legend.position="top")
p + theme(legend.position="bottom")
# Remove legend
p + theme(legend.position="none")
```
### Barplot with multiple groups


```{r}
df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))
head(df2)
```


```{r}
# Stacked barplot with multiple groups
ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(stat="identity") +
  theme_light()
# Use position=position_dodge()
ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
geom_bar(stat="identity", position=position_dodge())+
  theme_void()
```
































