---
title: "Day 9 R Packages and Data Reshaping in R"
author: "MAHESH DIVAKARAN"
date: "1/28/2021"
output: slidy_presentation
---

## R - Packages

R packages are a collection of R functions, complied code and sample data. They are stored under a directory called "library". All the packages available in R language are listed at R Packages. <https://cran.r-project.org/web/packages/available_packages_by_name.html>

### Available R Packages

Get library locations containing R packages

```{r}
.libPaths()
```
### List of all the packages installed

```{r}
library()

```


## All packages currently loaded in the R environment

```{r}
search()
```


## Install a New Package

There are two ways to add new R packages. One is installing directly from the CRAN and another is downloading the package to your local system and installing it manually. \ 

**1. Install directly from CRAN** \ 

`install.packages("Package Name")`

```{r}
# Install the package named "XML".
install.packages("XML")

```

**2. Install package manually** \

`install.packages(file_name_with_path, repos = NULL, type = "source")`

```{r}
# Install the package named "XML"
#install.packages("E:/XML_3.98-1.3.zip", repos = NULL, type = "source")
```

## Load Package to Library

A package is loaded using the following command −

`library("package Name", lib.loc = "path to library")`

```{r}
library(XML)
library(dichromat)

```




# R - Data Reshaping

## 1. Joining Columns and Rows in a Data Frame

We can join multiple vectors to create a data frame using the `cbind()` function. Also we can merge two data frames using `rbind()` function.

```{r}
# Create vector objects.
city <- c("Tampa","Seattle","Hartford","Denver")
state <- c("FL","WA","CT","CO")
zipcode <- c(33602,98104,06161,80294)

# Combine above three vectors into one data frame.
addresses <- cbind(city,state,zipcode)

addresses


# Create another data frame with similar columns
new_address <- data.frame(
   city = c("Lowry","Charlotte"),
   state = c("CO","FL"),
   zipcode = c("80230","33949"),
   stringsAsFactors = FALSE
)

new_address

# Combine rows form both the data frames.
all.addresses <- rbind(addresses,new_address)

all.addresses
```


## 2. Merging Data Frames

We can merge two data frames by using the `merge()` function.


```{r}
library(MASS)
Pima.te
Pima.tr

nrow(Pima.te)
nrow(Pima.tr)
```


```{r}
merged.Pima <- merge(x = Pima.te, y = Pima.tr,
   by.x = c("bp", "bmi"),
   by.y = c("bp", "bmi"),
)
merged.Pima
NROW(merged.Pima)

```


## Melting and Casting

One of the most interesting aspects of R programming is about changing the shape of the data in multiple steps to get a desired shape. The functions used to do this are called melt() and cast().

```{r}
library(MASS)
print(ships)
```

**Melt the Data**


Now we melt the data to organize it, converting all columns other than type and year into multiple rows.

```{r}
molten.ships <- melt(ships, id = c("type","year"))
print(molten.ships)
```

**Cast the Molten Data**

Cast the molten data into a new form where the aggregate of each type of ship for each year is created. It is done using the `cast()` function.

```{r}
recasted.ship <- cast(molten.ships, type+year~variable,sum)
print(recasted.ship)
```








