---
title: "Day 8 - Functions in R"
author: "MAHESH DIVAKARAN"
date: "1/27/2021"
output: pdf_document
---

# Functions in R

## Moving from Scripts to R Function

R function provides two major advantages over the script:

* Functions can work with any input. You can provide diverse input data to the functions.
* The output of the function is an object that allows you to work with the result.

```{r}
source('percentage_fn.R')
```

## Transforming the Script into R Function

**Stracture of the function**

function_name = `function`(_input_variable(s)){
staements
`return`(_output)
}

```{r}
# Creating the function

percentage_fn <- function(frac){
  percentage <- round(frac * 100, digits = 1)
  out <- paste(percentage, "%", sep = " ")
  return(out)
}
```


```{r}
#Calling percentage_fn
percentage_fn(0.3654)
```
```{r}
#Calling percentage_fn
percentage_fn(.00869)
```
### One more example


```{r}
basic_math <- function(a,b){
  add = a + b
  sub = a - b
  mul = a * b
  div = a / b
  print(paste("The sum of a and b is ", add ))
  print(paste("The difference of a and b is ",sub))
  print(paste("The product of a and b is ", mul))
  print(paste("The fraction of a and b is",round(div,3)))

}
```

### Run the basic_math() function
```{r}
basic_math(4,6)
```

```{r}
basic_math(854,966)
```

## R Vector Functions

`rep()`, `seq()`, `all()`, `any()`, `c()`

```{r}
# 1. R rep() Function

rep(c(0, 0, 7), times = 4)

rep(c(2, 4, 2), each = 2)

# 2. R seq() Function

seq(from = 4.5, to = 3.0, by = -0.5)

seq(1,16)

seq(10,100,5)

# 3. R any() Function

a = seq(10,100,5)
a
any( a > 100)

# 4. R all() Function

all(a > 0)


```

## R Numeric Functions

Function                          Description \

*abs(x)                              absolute value \
*ceiling(x)                          ceiling(3.475) is 4 \
*sqrt(x)                             square root \
*floor(x)                            floor(3.475) is 3 \ 
*log(x)                              natural logarithm \
*trunc(x)                            trunc(5.99) is 5 \
*round(x, digits=n)                  round(3.475, digit=2) is 3.48 \
*log10(x)                            common logarithm \
*signif(x, digits=n)                 signif(3.475, digit=2) is 3.5 \
*exp(x)                              e^x \
*cos(x), sin(x), tan(x)              also acos(x), cosh(x), acosh(zx) etc \


## Character Functions in R

```{r}
# 1. grep(pattern, x , ignore.case=FALSE, fixed=FALSE)

grep("A", c("b", "A", "c"), fixed = TRUE)

grep("an",c("an", "in", "and","back"),fixed = FALSE)

# 2. substr(x, start=n1, stop=n2)
string = "abcdef"
substr(string,2,4)

substr(string,2,4) <- "2222"
string

# 3. strsplit

strsplit("abc", "")

# 4. sub(pattern, replacement, x, ignore.case =FALSE, fixed=FALSE)

df<-"R is a collaborative project with many contributors"
#replaces the string 
sub('R','R language',df)

# 5. toupper(string)

toupper(df)

# 6. tolower(string)

tolower(df)


# 7. paste(…, sep=””)

paste("string", 1:3, sep = "")
paste(df,c(1,2,3,4))
```


