---
title: "Day 11 - R Datasets"
author: "MAHESH DIVAKARAN"
date: "2/1/2021"
output: html_document
---

## Importing Data to R



## Input as CSV File

*The csv file is a text file in which the values in the columns are separated by a comma. Let's consider the following data present in the file named input.csv.*

id,name,salary,start_date,dept\
1,Rick,623.3,2012-01-01,IT\
2,Dan,515.2,2013-09-23,Operations\
3,Michelle,611,2014-11-15,IT \
4,Ryan,729,2014-05-11,HR \
5,Gary,843.25,2015-03-27,Finance \
6,Nina,578,2013-05-21,IT \
7,Simon,632.8,2013-07-30,Operations \
8,Guru,722.5,2014-06-17,Finance


# Reading a CSV File

`read.csv()` function to read a CSV file available in your current working directory

```{r}
data <- read.csv("input.csv")
data
```
**Importing data from different Location with more import options**

`read.csv(file, header = TRUE, sep = ",", quote = "\"", dec = ".", fill = TRUE, comment.char = "",  na.strings = "NA", strip.white = TRUE, stringsAsFactors = FALSE)`



```{r}
data1 = read.csv("/cloud/project/Day 11/input1.csv",  sep = ",", header = T, na.strings = "NA", stringsAsFactors = T,dec = ".")

data1

```

```{r}
class(data$dept)

class(data1$dept)
```

**Importing data from web**
```{r}
#pizza_country_1990_2012.csv
data_web = read.csv("https://raw.githubusercontent.com/poldham/opensource-patent-analytics/master/2_datasets/infographic/pizza_country_1990_2012.csv",sep = ",", na.strings = "NA", strip.white = TRUE, stringsAsFactors = FALSE)

#data_web

#To see top 6 rows
head(data_web)
#To see bottom 6 rows
tail(data_web)
```











