---
title: "Day 9 R Packages"
author: "MAHESH DIVAKARAN"
date: "1/28/2021"
output: slidy_presentation
---

## R - Packages

R packages are a collection of R functions, complied code and sample data. They are stored under a directory called "library". All the packages available in R language are listed at R Packages. <https://cran.r-project.org/web/packages/available_packages_by_name.html>

### Available R Packages

Get library locations containing R packages

```{r}
.libPaths()
```
### List of all the packages installed

```{r}
library()

```


## All packages currently loaded in the R environment

```{r}
search()
```


## Install a New Package

There are two ways to add new R packages. One is installing directly from the CRAN and another is downloading the package to your local system and installing it manually. \ 

**1. Install directly from CRAN** \ 

`install.packages("Package Name")`

```{r}
# Install the package named "XML".
install.packages("XML")

```

**2. Install package manually** \

`install.packages(file_name_with_path, repos = NULL, type = "source")`

```{r}
# Install the package named "XML"
#install.packages("E:/XML_3.98-1.3.zip", repos = NULL, type = "source")
```

## Load Package to Library

A package is loaded using the following command −

`library("package Name", lib.loc = "path to library")`

```{r}
library(XML)
library(dichromat)

```








