---
title: "Day 7 - Control Structures in R"
author: "MAHESH DIVAKARAN"
date: "1/26/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Control Structures in R - Loops in R

To control the execution of the expressions flow in R we use the control structures. There are eight types of control structures in R. 

* if
* if-else
* for
* nested loops
* while
* repeat and break
* next
* return

## 1. if Condition in R
To carried out a task/command only `if` this condition is returned as *TRUE*.

**Syntax**

`if (test_expression) {statement}`


```{r, echo=TRUE}
#Example
x = sample(x = 1:100, size = 1,replace = F)
x

if(x %% 2 == 0 ){
  print(paste(x, " is an even number."))
}

```

## 1. if Condition in R

```{r, echo=TRUE}
# Example 2 
values <- 1:10                                        
if (sample(values,1) <= 10)                               
print(paste(values, "is less than or equal to 10"))
```

## 2. if-else Condition in R

**Syntax**

`if (test_expression) {
statement
} else {
statement
}
`

```{r, echo=TRUE}
#Example
x

if(x %% 2 == 0 ){
  print(paste(x, " is an even number."))
}else{
  print(paste(x, " is an odd number."))
}
```

## 2. if-else Condition in R

```{r,echo=TRUE}
#example 2
val1 = 10           #Creating our first variable val1
val2 = 5            #Creating second variable val2
if (val1 > val2){   #Executing Conditional Statement based on the comparison 
  print("Value 1 is greater than Value 2")
} else if (val1 < val2){
  print("Value 1 is less than Value 2")
}
```

## 3. for Loop in R

A *loop* is a sequence of instructions that is repeated until a certain condition is reached.  for, while and repeat, with the additional clauses break and next are used to construct loops.
`for` loop is executed until a known number of times(finite).

**Syntax**

`for(test_expression){statement}`

```{r, echo=TRUE}
# Example
values <- c(1,2,3,4,5) 
for(id in 1:5){ 
  print(values[id])
}
```
## 4. Nested Loop in R

A loop inside another loop is defined as **nested loop**.

```{r,echo=TRUE}
## Create a matrix

mat <- matrix(1:10, 2)
mat
```
## 4. Nested Loop in R
```{r,echo=TRUE}
#Example
for (id1 in seq(nrow(mat))) {
  for (id2 in seq(ncol(mat))) {
    print(mat[id1, id2])
  }
}
```

## 5. while Loop in R

In R programming, while loops are used to loop until a specific condition is met.

**Syntax**

`while (test_expression){statement}`

```{r,echo=TRUE}
i <- 1
while (i < 6) {
print(i)
i = i+1
}
```

## 6. repeat and break Statement in R

We use `break` statement inside a loop (repeat, for, while) to stop the iterations and flow the control outside of the loop. 
A `repeat` loop is used to iterate over a block of code, multiple numbers of times. 

**Syntax**

`repeat {`
simulations; generate some value have an expectation if within some range, then exit the loop
`if ((value - expectation) <= threshold) {`
`break`
`}`
`}`
```{r, echo=TRUE}
#Example
val <- 5
repeat {
  print(val)
  val <- val+1
  if (val == 10){
    break
  }
}
```
## 6. repeat and break Statement in R

```{r, echo=TRUE}
# Example of Break Statement in R
values = 1:10
for (id in values){
  if (id == 4){
    break
  }
  print(id)
}
```

## 7. next Statement in R

`next` jumps to the next cycle without completing a particular iteration.

```{r,echo=TRUE}
# Example
x = 1: 4
for (i in x) {
if (i == 2) {
next
}
print(i)
}
```

## 8. return Statement in R

`return()` function  do processing and return back the result.

**Syntax**
`return(expression)`

```{r, echo=T}
check <- function(x) {
if (x > 0) {
result <- "Positive"
} else if (x < 0) {
result <- "Negative"
} else {
result <- "Zero"
}
return(result)
}

check(7852)
```

## Thank You!!




