---
title: "R for Data Analysis - Day 6"
author: "MAHESH DIVAKARAN"
date: "1/25/2021"
output: html_document
---


# Welcome to R Markdown!!!


```{r}
print("Welcome to R Markdown!!!")
```

## Introduction to Data Structures in R

There are many types of R-objects. The popularly used ones are:

Vector
Matrix
Array
Lists
Data Frames

### 1. R Vector

A vector is a sequence of elements that share the same data type. 

R vector comes in two parts: `Atomic vectors` and `Lists`.

```{r}
vector() # an empty 'logical' (the default) vector

vector("character", length = 5) # a vector of mode 'character' with 5 elements

character(5) # the same thing, but using the constructor directly

numeric(5)   # a numeric vector with 5 elements

logical(5)   # a logical vector with 5 elements

# You can also create vectors by directly specifying their content. R will then guess the appropriate mode of storage for the vector. 

a = c(1,2,3,4,5,6,7)

a

typeof(a)

x <- c(1, 2, 3)

x
typeof(x)

x1 <- c(1L, 2L, 3L)
x1 

typeof(x1)


# coerce to the integer type using as.integer()

x2 =  as.integer(x)
x2
typeof(x2)

# Using TRUE and FALSE will create a vector of mode logical

y <- c(TRUE, TRUE, FALSE, FALSE)

y

typeof(y)

# Quoted text will create a vector of mode character

z <- c("Sarah", "Tracy", "Jon")

zz <- c("Sarah", "Tracy", "Jon","Sarah", "Tracy", "Jon","Sarah", "Tracy", "Jon","Sarah", "Tracy", "Jon")

# Examining Vectors

# The functions typeof(), length(), class() and str() provide useful information about your vectors and R objects in general.

typeof(z)
length(z)
class(z)
str(z)
str(zz)


mult = c(1,2,3,"Sarah", "Tracy", "Jon",T,FALSE,1+9i)
typeof(mult)
class(mult)

# Adding Elements in a Vector
#The function c() (for combine) can also be used to add elements to a vector.

z <- c(z, "Annette", "Biju")
z

z <- c("Greg", z)



# Vectors from a Sequence of Numbers

series <- 1:100

series1 = seq(100)


ser1 = seq(from = 1, to = 10, by = 0.1)

ser2 = seq(1,10,0.1)

ser1 == ser2


# Missing Data

# R supports missing data in vectors. They are represented as NA (Not Available) and can be used for all the vector types covered in this lesson:

x <- c(0.5, NA, 0.7)

#The function is.na() indicates the elements of the vectors that represent missing data, and the function anyNA() returns TRUE if the vector contains any missing values


is.na(x)

anyNA(x)

x1 <- c(TRUE, FALSE, NA)

is.na(x1)

anyNA(x1)


#Other Special Values

#Inf is infinity. You can have either positive or negative infinity.

1/0
-1/0

0/0


```



### 2. R List

R list is the object which contains elements of different types – *like strings, numbers, vectors and another list inside it*.

```{r}


# Create a list containing strings, numbers, vectors and a logical values

list_data <- list("Red", "Green", c(21,32,11), TRUE, 51.23, 119.1)
list_data

#Accessing List Elements

list_data[1]
list_data[length(list_data)]

list1 = list(1,2,3,4,5,6,7,8,9,10)

list1[8]

list1[8] = 80
list1


# Create two lists.
list1 <- list(1,2,3)
list2 <- list("Sun","Mon","Tue")
list3 <- list(1:3)

# Merge the two lists.
merged.list <- c(list1,list2,list3)

# Print the merged list.
print(merged.list)


# Converting List to Vector

#we use the unlist() function to convert list to vector.

# Create lists.
list1 <- list(1:5)
print(list1)
list1
list1[[1]][1]
list1[[1]][5]

list2 <-list(10:14)
print(list2)


# Convert the lists to vectors.
v1 <- unlist(list1)
v2 <- unlist(list2)

print(v1)
print(v2)

length(v1)
length(v2)


add = v1 + v2
sub = v1 - v2
mult = v1 * v2
div = v1 / v2
```


### 3. R Matrix

In a matrix, numbers are arranged in a fixed number of rows and columns and usually, the numbers are the real numbers. *All the data elements must share a common basic type.*


`Funtion - matrix()`

```{r}
matrix()
```

```{r}
mat <- matrix (
c(2 , 4, 3, 1, 5, 7),       # the data elements
nrow =2,                   # no. of rows
ncol =3,                   # no. of columns
byrow = TRUE)

mat

# An element at the mth row and nth column 
mat[2, 3]

# To extract only the mth row of our matrix ‘mat’
mat[2,]

# To extract only the nth column of our matrix ‘mat’
mat[,3]

```

Creating R matrix through the arrangement of elements sequentially by row
```{r}
arrang_row <- matrix(c(4:15), nrow = 4, byrow = TRUE) #Creating our matrix and arranging it by row
print(arrang_row) #Printing our arranged matrix
```
Creating R matrix by arranging elements sequentially by column

```{r}
arrang_col <- matrix(c(4:15), ncol = 4, byrow = FALSE)       #Creating our matrix and arranging it by column

print(arrang_col)    #Printing our arranged matrix 
```

#### Defining names of columns and rows in a matrix

```{r}
rows = c("row1", "row2", "row3", "row4")     #Creating our character vector of row names

cols = c("colm1", "colm2", "colm3")          #Creating our character vector of column names

mat <- matrix(c(4:15), nrow = 4, byrow = TRUE, dimnames = list(rows, cols) )   
#creating our matrix mat and assigning our vectors to dimnames

print(mat)                    #Printing our matrix
```

#### Modify Matrix in R

Assign a Single Element

```{r}
mat
mat[2,2] = 20 # replae single element
mat
mat[,2] = c(2,2,2,2) # Replace column
mat
mat[3,] = c(2,2,2) # Replace row
mat
```

Use of Relation Operators

```{r}
mat[mat == 2] <- 0     #Replacing elements that are equal to 2 with 0
mat

mat[mat < 10] <- 0       #Replacing elements that are less with 10 with 0
mat


```


Addition of Rows and Columns

```{r}
new_mat = matrix(1:12, nrow = 3, ncol = 3)  # Creating a new matrix
new_mat

# add a column to our matrix ‘new_mat’ using cbind() function

new_mat = cbind(new_mat, c(1,2,3))
new_mat

# add a row using the rbind() function

new_mat = rbind(new_mat, c(1,2,3,4))
new_mat

# dimension of the matrix ‘new_mat’ using the dim() function
dim(new_mat)

# modify the dimension of the matrix 
dim(new_mat) <- c(2,8)
new_mat

# transpose of the matrix using the t() function
t(new_mat)
```

### R Matrix Operations
1. Addition (+)

```{r}
mat1 <- matrix(data = 1:8, nrow = 4, ncol = 4) #Creating our first matrix mat1 
mat1
mat2 <- matrix(data = 1:16, nrow = 4, ncol = 4) #Creating our second matrix mat2
mat2

sum <- mat1 + mat2 #Adding our two matrices
print(sum) #Printing the sum
```

2. Subtraction (-)

```{r}
sub <- mat1 - mat2 #Adding our two matrices
print(sub) #Printing the sum
```

3. Matrix Multiplication (By Constant)

```{r}
prod <- mat1*4 #Multiplying matrix mat1 with constant value 4
print(prod) #Printing the product
```

4. Multiplication (*)

```{r}
prod <- mat1*mat2 #Multiplying matrix mat1 with mat2
print(prod) #Printing the product
```

5. Division (/)

```{r}
div <- mat1/mat2 #Division of mat1 and mat2 
print(div) #Printing the division
```


## R Array 

R arrays are the data objects which can store data in more than two dimensions. In arrays, data is stored in the form of matrices, rows, and columns. 

An array is created using the `array()` function.






























