---
title: "Linear Regression with Boston Housing Dataset"
author: "Mahesh_Divakaran"
date: "24 November 2018"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
```
#Linear Regression 
First, lets Install and load the requried packages. 

```{r eval=FALSE}
install.packages('readr')
install.packages('ggplot2')
install.packages('mlbench')
install.packages('corrplot')
install.packages('Amelia')
install.packages('caret')
install.packages('plotly')
install.packages('caTools')
install.packages('reshape2')
install.packages('dplyr')
```

```{r message=FALSE,warning=FALSE,error=FALSE}

library(readr)
library(ggplot2)
library(corrplot)
library(mlbench)
library(Amelia)
library(plotly)
library(reshape2)
library(caret)
library(caTools)
library(dplyr)
```
##Boston Housing Dataset
Housing data contains 506 census tracts of Boston from the 1970 census. The dataframe `BostonHousing` contains the original data by Harrison and Rubinfeld (1979), the dataframe BostonHousing2 the corrected version with additional spatial information.

You can include this data by installing `mlbench` library.The data has following features, `medv` being the target variable:

 * crim   - per capita crime rate by town
 * zn     - proportion of residential land zoned for lots over 25,000 sq.ft
 * indus	- proportion of non-retail business acres per town
 * chas	  - Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)
 * nox	  - nitric oxides concentration (parts per 10 million)
 * rm	    - average number of rooms per dwelling
 * age	  - proportion of owner-occupied units built prior to 1940
 * dis	  - weighted distances to five Boston employment centres
 * rad	  - index of accessibility to radial highways
 * tax	  - full-value property-tax rate per USD 10,000
 * ptratio- pupil-teacher ratio by town
 * b	1000(B - 0.63)^2, where B is the proportion of blacks by town
 * lstat  - percentage of lower status of the population
 * medv	  - median value of owner-occupied homes in USD 1000's

 
##Data 
**Load the `BostonHousing` data and assign it to the variable `housing` . **
 
```{r message=FALSE,warning=FALSE}
data(BostonHousing)
housing <- BostonHousing
str(housing)
```

Lets examine the `head` of the `housing` dataframe using `head()`.
```{r message=FALSE,warning=FALSE}
head(housing)
```

```{r message=FALSE,warning=FALSE}
summary(housing)
```

##Data Cleaning
Next we have to clean this data.There are many ways to do this. I will be using `missmap()` from `Amelia` package. 

### Check for any NA's in the dataframe. 

```{r message=FALSE,warning=FALSE}
missmap(housing,col=c('yellow','black'),y.at=1,y.labels='',legend=TRUE)

```

The above plot clearly shows that the data is free from NA's. 

##Exploratory Data Analysis

Let's use ggplot2,corrplot and plotly to explore the data a bit. 

###Visualizations {.tabset}

####Correlation 

**Correlation and CorrPlots**

From Wikipedia, correlation is defined as:

In statistics, **dependence or association** is any statistical relationship, whether causal or not, between two random variables or two sets of data. 
Correlation is any of a broad class of statistical relationships involving dependence, though in common usage it most often refers to the extent to 
which two variables have a linear relationship with each other. 

**Correlation** plots are a great way of exploring data and seeing if there are any interaction terms.

```{r message=FALSE,warning=FALSE,error=FALSE,fig.align='center'}

corrplot(cor(select(housing,-chas)))
```

`medv` decreases with increase in  `crim` (*medium*), `indus` (*High*),`nox`(*low*),`age`(*low*),`rad`(*low*),`tax`(*low*),`ptratio`(*high*), `lstat` 
(*High*) and increases with increase in `zn`(*low*),`rm`(*High*).

####medv Density Plot using ggplot2

```{r message=FALSE,warning=FALSE,error=FALSE,fig.align='center'}
#visualizing the distribution of the target variable 

housing %>% 
  ggplot(aes(medv)) +
  stat_density() + 
  theme_bw()
```
The above visualizations reveal that peak densities of `medv` are in between 15 and 30. 

####medv Density using plotly
```{r message=FALSE,warning=FALSE,error=FALSE,fig.align='center'}
ggplotly(housing %>% 
  ggplot(aes(medv)) +
  stat_density() + 
  theme_bw())
```

The above visualizations reveal that peak densities of `medv` are in between 15 and 30. 

####`medv`
Let's see the effect of the variables in the dataframe on `medv`. 
```{r message=FALSE,warning=FALSE,error=FALSE,fig.align='center'}
housing %>%
  select(c(crim, rm, age, rad, tax, lstat, medv,indus,nox,ptratio,zn)) %>%
  melt(id.vars = "medv") %>%
  ggplot(aes(x = value, y = medv, colour = variable)) +
  geom_point(alpha = 0.7) +
  stat_smooth(aes(colour = "black")) +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  labs(x = "Variable Value", y = "Median House Price ($1000s)") +
  theme_minimal()
```

The results from the above graph are in correlation with the corrplot.

##Model Building & Prediction 

###General Form 

The General Linear regression model in R : 

Univariate Model :
$model <- lm(y\sim x,data)$

Multivariate Model  :
$model <- lm(y \sim.,data)$

###Train and Test Data

Lets split the data into train and test data using `caTools` library.

```{r message=FALSE,warning=FALSE,error=FALSE}
#set a seed 
set.seed(123)

#Split the data , `split()` assigns a booleans to a new column based on the SplitRatio specified. 

split <- sample.split(housing,SplitRatio =0.75)


train <- subset(housing,split==TRUE)
test <- subset(housing,split==FALSE)

# train <- select(train,-b)
# test <- select(test,-b)

```

###Training our Model 
Lets build our model considering that `crim`,`rm`,`tax`,`lstat` as the major influencers on the target variable.

```{r error=FALSE,warning=FALSE,message=FALSE}
model <- lm(medv ~ crim + rm + tax + lstat , data = train)
summary(model)
```

###Visualizing our Model
Lets visualize our linear regression model by plotting the residuals. The difference between the observed value of the **dependent variable** `(y)` and
the **predicted value** `(y)` is called the **residual** `(e)`.

```{r message=FALSE,warning=FALSE}

res <- residuals(model)

# Convert residuals to a DataFrame 

res <- as.data.frame(res)
```

```{r message=FALSE,warning=FALSE,fig.align='center'}
ggplot(res,aes(res)) +  geom_histogram(fill='blue',alpha=0.5)

plot(model)
```

###Predictions

Let's test our model by predicting on our testing dataset. 

```{r message=FALSE,warning=FALSE,fig.align='center'}
test$predicted.medv <- predict(model,test)

pl1 <-test %>% 
  ggplot(aes(medv,predicted.medv)) +
  geom_point(alpha=0.5) + 
  stat_smooth(aes(colour='black')) +
  xlab('Actual value of medv') +
  ylab('Predicted value of medv')+
  theme_bw()

ggplotly(pl1)
```

Lets evaluate our model using Root Mean Square Error, a standardized measure of how off we were with our predicted values.

###Assessing our Model

```{r message=FALSE,warning=FALSE}
error <- test$medv-test$predicted.medv
rmse <- sqrt(mean(error)^2)
```


The Root Mean Square Error (RMSE) for our Model is `r rmse` and the Results can be further improved using feature extraction and rebuilding,training 
the model.
