---
title: "Regression Analysis"
author: "MAHESH DIVAKARAN"
date: "2/15/2021"
output: html_document
---

### Loading Libraries
```{r}
library(fBasics)  #jarque-bera 
library(lmtest)   # Durbin-Watson test
```

## Reading the dataset

The data we will use, one year of marketing spend and company sales by month.

```{r}
dataset = read.csv("/cloud/project/Data/budget.csv", header = T, colClasses = c("numeric", "numeric", "numeric"))
dataset
```

```{r}
str(dataset)

```

```{r}
#Plotting the data'

plot(dataset$Spend,dataset$Sales)

plot(dataset$Month,dataset$Spend)
plot(dataset$Month,dataset$Sales)
```

## Linear Regression Using lm()

**Simple Linear Regression Using `lm()`**

The Simple Linear Regression equation,  y = β0 + β1X 

```{r}
simple.fit = lm(Sales ~ Spend, data=dataset)
summary(simple.fit)
```

```{r}
layout(matrix(c(1,1,2,3),2,2,byrow=T))
#Spend x Residuals Plot
plot(simple.fit$resid~dataset$Spend[order(dataset$Spend)],
 main="Spend x Residuals\nfor Simple Regression",
 xlab="Marketing Spend", ylab="Residuals")
abline(h=0,lty=2)
#Histogram of Residuals
hist(simple.fit$resid, main="Histogram of Residuals",
 ylab="Residuals")
#Q-Q Plot
qqnorm(simple.fit$resid)
qqline(simple.fit$resid)
```


The histogram and QQ-plot are the ways to visually evaluate if the residual fit a normal distribution. 

```{r}
#Null Hypothesis: Skewness and Kurtosis are equal to zero
jarqueberaTest(simple.fit$resid) #Test residuals for normality
```

With a p value of 0.6195, we fail to reject the null hypothesis that the skewness and kurtosis of residuals are statistically equal to zero.

```{r}
# The Null hypothesis of the Durbin-Watson test is that the errors are serially UNcorrelated.
dwtest(simple.fit) #Test for independence of residuals  
```



# Multiple Linear Regression


###The Libraries required
```{r}
library(datarium)
library(caTools)
library(ggplot2)
library(GGally)
```

# The data corresponding to youtube, Facebook, newspaper and sales the data size comes out to be 200 x 4.
```{r}
#Loading The Data
data("marketing", package = "datarium")
data_size = dim(marketing)
data_size
```
```{r}
# Understanding the data
head(marketing,5)
summary(marketing)
```

```{r}
#Pairwise plotting technique 1
plot(marketing, col="purple", main="Plotting Pairs Against Each Other")
```

```{r}
#Pairwise plotting technique 2
ggpairs(marketing)
```

### Preparing the data

```{r}
#Global Variable
splitRatio = 0.80

#Splitting The Data
set.seed(101)# Set Seed so that same sample can be reproduced in future also

#Now Selecting 80% of data as sample from total 'n' rows of the data
sample = sample.split(marketing$youtube, SplitRatio = splitRatio)
sample
dim(sample)
train = subset(marketing, sample == TRUE)
test = subset(marketing, sample == FALSE)
train_size = dim(train)
train_size
test_size = dim(test)
test_size
```


### Creating the model

sales ~ youtube + facebook + newspaper

```{r}
#creating the model
Model <- lm(sales ~ youtube + facebook + newspaper, data = marketing)
summary(Model)
```

#### Model: sales  = 3.526 + 0.045 * youtube + 0.188* facebook + 0.001 * newspaper

## Model Accuracy Analysis
```{r}
#Predicting
pred <- predict(Model, test)
numx <- data_size[1]*(1 - splitRatio)
x_axis <- seq(numx)
df <- data.frame(x_axis, pred,test$sales)
df
```

```{r}
#Plotting the predicted values against the actual values
g <- ggplot(df, aes(x=x_axis))
g <- g + geom_line(aes(y=pred, colour="Predicted"))
g <- g + geom_point(aes(x=x_axis, y=pred, colour="Predicted"))
g <- g + geom_line(aes(y=test$sales, colour="Actual"))
g <- g + geom_point(aes(x=x_axis, y=test$sales, colour="Actual"))
g <- g + scale_colour_manual("", values = c(Predicted="red", Actual="blue"))
g
```

```{r}
#Evaluation
original = test$sales
predicted = pred
d = original-predicted
mse = mean((d)^2)
mae = mean(abs(d))
rmse = sqrt(mse)
R2 = 1-(sum((d)^2)/sum((original-mean(original))^2))
cat(" MAE:", mae, "\n", "MSE:", mse, "\n", 
    "RMSE:", rmse, "\n", "R-squared:", R2)
```


```{r}
#Histogram of Residuals
hist(Model$resid, main="Histogram of Residuals",
 ylab="Residuals")
#Q-Q Plot
qqnorm(Model$resid)
qqline(Model$resid)
```























